### 2018-06-11 main.tex
first commit

\documentclass{article}

\usepackage{mystyle}
\usepackage{mycommands}

\title{Continuous Reinforcement Learning}

\author{
  Corentin Tallec, Léonard Blier
}

\begin{document}
\maketitle

\begin{abstract}
Abstract
\end{abstract}

\section{Introduction}
\label{sec:introduction}

\section{Definitions and Notations}
\label{sec:definitions}

\begin{definition}[Continuous Markov Reward Process]
  A continuous markov decision process $(X_{t}, R_{t})_{t\in\BR_{+}}$ is a continuous random process defined by 
  \begin{itemize}
  \item $\mathcal{X}$ is the state set 
  \item 
  \end{itemize}
  process$\mathcal{X}$ is the set of states, $T$ a transition function, $r$ a reward function
\end{definition}

\section{Ordre de grandeur Corentin}
\label{sec:ordre-de-grandeur}

\section{MCTS continu}
\label{sec:mcts-continu}

\section{Conclusion}
\label{sec:conclusion}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: