### 2018-10-18 paper/main.tex

\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}

\newcommand{\deltat}{{\delta t}}
\newcommand{\gauss}{\mathcal{N}}
\newcommand{\E}{\mathbb{E}}

\title{Scaling continuous reinforcement learning up}
\begin{document}
\maketitle
\begin{abstract}
	Most reinforcement learning algorithms are not resilient to a change of
	time discretization. Even worse, their time continuous limit is often
	ill defined.  While continuous reinforcement methods have been
	introduced in the past, they have been dismissed when scaling up from
	reinforcement learning to deep reinforcement learning. We introduce the
	first deep reinforcement learning that scales when faced with
	diminishing discretization steps.
\end{abstract}
\paragraph{Problem at hand}
Assume a stochastic continuous environment of the form
\begin{align}
	s_{t+\deltat} &= s_t + F(s_t, a_t) \deltat + \gauss(0, C(s_t, a_t)) \sqrt{\deltat}\\
	r(s_t, a_t) &= \tilde{r}(s_t, a_t) \deltat
\end{align}
with $F$ a deterministic drift, and $C$ a noise covariance matrix
associated to each $(s, a)$ pair. Our aim is to find, a policy $\pi$, that is a
(potentially stochastic) mapping from states to actions that maximizes the
average return, i.e. the expected average of rewards
\begin{align}
	R^\pi &= \lim\limits_{T\to+\infty}
	\frac{1}{T}
	\E_{s_0, a_0, \ldots, s_T, a_T\sim\pi}\left[
		\sum\limits_{t = 0}^T \tilde{r}(s_t, a_t)
	\right]
\end{align}
or the discounted sum of future rewards
\begin{align}
	R^\pi_\gamma &= \E_{s_0, a_0, \ldots, s_T, a_T \sim \pi}
	\left[
		\sum\limits_{t=0}^{+\infty}
		(\gamma^\deltat)^t \tilde{r}(s_t, a_t) 
	\right].
\end{align}
Further assume that for any policy, the environment admits an invariant
measure, that is that there exists a distribution on states $\rho^\pi$ such that for
any regular function $F$ and any infinite sequence $s_0, \ldots$ sampled using policy
$\pi$,
\begin{equation}
	\lim\limits_{T\to+\infty}
	\frac{1}{T}
	\sum\limits_{t=0}^T F(s_t) =
	\E_{s\sim \rho^\pi}\left[
		F(s)
	\right].
\end{equation}
\paragraph{Advantage updating}
\end{document}