### 2018-11-29 paper/learn_discontinuities.tex
Learning discontinuities for A

\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}

\newcommand{\E}{\mathbb{E}}
\newcommand{\bern}[1]{\mathcal{B}\left(#1\right)}
\newcommand{\gauss}[1]{\mathcal{N}\left(#1\right)}
\newcommand{\argmax}{\text{amax}}
\newcommand{\bigO}{O}

\begin{document}
Consider the advantage updating set of equations in continuous time
\begin{align}
	A^*(s, a) &= \max_a A^*(s,a) + \frac{r_t dt + \gamma^{dt} \E_{s_{t+dt}} V^*(s_{t+dt}) - V^*(s_t)}{dt}\\
	\max_a A^*(s, a) &= 0.
	\label{eq:adv_up}
\end{align}
On ergodic reversible environments, those equations are mostly well defined in
term of orders of magnitude, as $Q^*(s,a) - V^*(s) = \bigO(dt)$ for all $s, a$.
However, when facing highly non reversible environments, e.g. episodic
environments, action state pairs where the advantage between two different actions
is not of order $dt$ exist, i.e., there can exist $s_0, a_0$ such that
$Q^*(s_0, a_0) - V^*(s_0) = \bigO(dt)$. In such cases, $A^*(s_0, a_0)$ is infinite,
and it is doubtful that an algorithm learning parametric approximations of
$V$ and $A$ based on~\eqref{eq:adv_up} will properly converge.

To palliate this behaviour, when learning parametric approximations of $V^*$ and $A^*$,
$V_\theta$, $A_\phi$, one can parameterize $A_\phi$ as a mixture of gaussians
\begin{align}
	A_\phi(s, a) = \bern{\mu_\phi(s, a)dt} \frac{\gauss{\tilde{A}_\phi(s, a), 1}}{dt} + \bern{1 - \mu_\phi(s, a)dt} \gauss{\bar{A}_\phi(s, a), 1}
\end{align}
where $\bern{p}$ denotes independant bernoulli random variables of parameter $p$, and
$\gauss{\mu, \sigma}$ independant gaussian random variables of mean $\mu$, variance
$\sigma$.

Suppose the value of $V^*(s)$ and $V^*(s')$ given on a dataset of transitions.
We want to compute the likelihood of our model of $A$.
The likelihood of the model of $A$ on a single datapoint is given by the log likelihood
of the following mixture density
\begin{align}
	&\bern{\mu(s, a)\mu(s, \argmax_{a'} \tilde{A}(s, a')dt^2} \frac{\gauss{\tilde{A}(s, a) - \max_a \tilde{A}(s, a), 2}}{dt} + \nonumber\\
	&\bern{(1 - \mu(s, a)dt)\mu(s, \argmax_{a'} \tilde{A}(s, a')dt} \frac{\gauss{\bar{A}(s, a)dt - \max_a \tilde{A}(s, a), 1 + dt^2}}{dt} + \nonumber\\
	&\bern{(1 - \mu(s, a)dt)(1 - \mu(s, \argmax_{a'} \bar{A}(s, a')dt)} \gauss{\bar{A}(s, a) - \max_a \bar{A}(s, a), 2} + \nonumber\\
	&\bern{\mu(s, a)dt(1 - \mu(s, \argmax_{a'} \bar{A}(s, a')dt)} \frac{\gauss{\tilde{A}(s, a) - \max_a \tilde{A}(s, a), 1 + dt^2}}{dt}
\end{align}
from which a gradient based learning procedure can be derived for $A$.
\end{document}