### 2018-11-30 paper/baselines.tex

\documentclass[11pt,a4paper]{article}
\input{header.en.tex}

\begin{document}
Let us gather a few baselines for continuous-time RL, starting with the
most basic. The baselines below have no problem with the continuous-time
limit (provided $\gamma<1$).

All to be tested with the same exploration strategies etc. (I want to
test how much the problem lies with solving Bellman.)

\begin{enumerate}

\item Reinforce (online version). High variance. All generalization comes
from the policy. Therefore the policy had better have good features. Thus
the performance of Reinforce may improve significantly by providing
enhanced features to the policy, eg features coming from an auto-encoder
or world model. (I think this applies to all algorithms, actually, and
could be quite important.)

\item TD(1) then policy gradient or a variant of policy gradient.

\item TD(1) trained on a model of the reward, then some form of policy
gradient.

\item 4?
\end{enumerate}

Baselines that should not (in principle) pass to the limit:
\begin{itemize}
\item All fashionable deep RL algorithms, DQN, TRPO, PPO,...
\item DPG when applicable
\end{itemize}

\end{document}