### 2018-12-05 paper/icml_drau/icml_drau.tex

%%%%%%%% ICML 2018 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{amsmath, amssymb}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables
\usepackage{xcolor}

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
\usepackage{hyperref}

\input{macros.tex}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2018}

% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{icml2018}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Deep Relativistic Advantage Updating}

\begin{document}

\twocolumn[
\icmltitle{There are no small Advantages:
	Reinforcement learning in near continuous time
	with Deep RelAtive Mixture of Advantages
}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2018
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
%\icmlauthor{Aeiau Zzzz}{equal,to}
%\icmlauthor{Bauiu C.~Yyyy}{equal,to,goo}
%\icmlauthor{Cieua Vvvvv}{goo}
%\icmlauthor{Iaesut Saoeu}{ed}
%\icmlauthor{Fiuea Rrrr}{to}
%\icmlauthor{Tateu H.~Yasehe}{ed,to,goo}
%\icmlauthor{Aaoeu Iasoh}{goo}
%\icmlauthor{Buiui Eueu}{ed}
%\icmlauthor{Aeuia Zzzz}{ed}
%\icmlauthor{Bieea C.~Yyyy}{to,goo}
%\icmlauthor{Teoau Xxxx}{ed}
%\icmlauthor{Eee Pppp}{ed}
\end{icmlauthorlist}

%\icmlaffiliation{to}{Department of Computation, University of Torontoland, Torontoland, Canada}
%\icmlaffiliation{goo}{Googol ShallowMind, New London, Michigan, USA}
%\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}
%
%\icmlcorrespondingauthor{Cieua Vvvvv}{c.vvvvv@googol.com}
%\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Reinforcement learning}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}

\end{abstract}

\section{Introduction}
\label{sec:intro}

\section{Related work}
\label{sec:related}

\section{Continuous reinforcement learning and scaling problems}
\label{sec:continous}
In the following section, a restrictive framework for near continuous
time reinforcement learning is introduced, and the main challenges of this
setup are explained. \RKC{This is not what we want to write: we want to explain
why current reinforcement learning practices are ill suited for this framework}

\subsection{Framework}
Consider a fully observable \emph{Markov Decision Process} defined by the following
transitions and rewards
\begin{align}
	s_{t + \deltat} &= s_t + F(s_t, a_t) \deltat + \Sigma(s_t, a_t) \deltabt\\
	r_{t} &= \reward(s_t, a_t) \deltat + \Sigma_r(s_t, a_t) \deltabt^r
\end{align}
where the $s$'s are $n$-dimensional vectors, $F(s_t, a_t)$ is a $n$-dimensional
drift coeficient, $\Sigma(s_t, a_t)$ is a $n \times n$ diffusion matrix,
$\deltat$ is a small but finite discretization step, and $\deltabt$ is a sample
from $\gauss(0, \deltat I)$.

This framework is quite general, and most discretized deterministic continuous
control environments fall in this framework, with $\Sigma = 0$ and $\Sigma_r =
0$, as well as many stochastic control environments.

A deterministic policy $\pi$ is defined as a deterministic mapping from state
space to action space\footnote{We restrict ourselves to deterministic policies.
	In continuous time, defining stochastic policies with well defined
	state-value function is non trivial. In the discrete case, the
distribution of actions of a policy given a state only depends on the previous
state, and not on previous actions. In continuous time, such policies are akin
to white noise, and do not define stochastic processes.}.

Under a deterministic policy $\pi$, one can define the \emph{state value function} $V^\pi(s)$
and \emph{state action value function} $Q^\pi(s, a)$ as
\begin{align}

\end{align}

\section{Multi-chain continuous environments and biscale advantage}
\label{sec:multi-chain}

\section{Deep RelAtivistic Mixture of Advantage learning}
\label{sec:algo}

\section{Experiments}
\label{sec:exp}

\section{Conclusion}
\label{sec:concl}

\bibliography{icml_drau.bib}
\bibliographystyle{icml2018}
\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018. It was modified from a version from Dan Roy in
% 2017, which was based on a version from Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.