### 2018-12-10 paper/tangotd.tex

\documentclass[11pt,a4paper]{article}
\input{header.en.tex}

\newcommand{\Dir}{\mathcal{D}}

\begin{document}

\section{Recovering known convergence results for TD, and a few ideas}

We consider the Bellman equation for a policy $\pi$ in a finite MDP with
expected reward $R$ and transition probability matrix $P$ ($P$ includes $\pi$ and
the environment response). The Bellman equation is
\begin{equation}
(\Id-P)V=R
\end{equation}

We assume we can sample transitions $(s,a,r,s')$ according to some
distribution $\rho$ on $s$. In the on-policy case, $\rho$ will be the
stationary distribution. Typically $a$ has to be taken from $\pi$ (or a
distribution close to $\pi$ after resampling/by using only those $a$
sampled from $\pi$; this covers $\eps$-greedy) so that we do get
transitions from
$P$. (Alternatively one could work with pairs $(s,a)$ TODO.) Thus, we
have access to unbiased samples of $\rho P$ and $\rho R$:
\begin{equation}
\rho P=\E[\1_{s}\transp{\1_{s'}}\,],\qquad \rho R=\E[r(s,a,s')\1_{s}]
\end{equation}
and thus also to samples of $\rho(\Id-P)$. Thus, rewriting the equation
as
\begin{equation}
\rho(\Id-P)V=\rho R
\end{equation}
(equivalent if $\rho$ is positive), both sides are \emph{samplable}, and we can try to solve the above by
stochastic gradient-like methods.

For instance, if we parameterize $V$ by some parameter $\theta$, we can apply
parametric TD. The expected TD step is
\begin{equation}
\theta\gets \theta -\eta \,\transp{\Phi}\rho (V-PV-R)
\end{equation}
where 
\begin{equation}
\Phi\deq \frac{\partial V}{\partial \theta}
\end{equation}
is the Jacobian matrix of $V$ with respect to its parameter. The linear
case $V=\Phi \theta$ corresponds to constant $\Phi$. The tabular case is
$\Phi=\Id$.

A fixed point of TD solves
\begin{equation}
\transp{\Phi}\rho(\Id-P)V=\transp{\Phi}\rho R
\end{equation}
and if the system is overparameterized ($\Phi$ invertible) and
well-sampled ($\rho>0$) this implies $(\Id-P)V=R$. All terms in this
equation are \emph{samplable}: we can sample from $\rho R$ and from
$\rho(\Id-P)$.

Thus, from now on we are going to focus on solving
\begin{equation}
\transp{\Phi}\rho (\Id-P)V=\transp{\Phi}\rho R
\end{equation}
using various SGD-like solvers.

We now switch to the linear case and will de-linearize later. This
becomes
\begin{equation}
\transp{\Phi}\rho (\Id-P)\Phi \theta=\transp{\Phi}\rho R
\end{equation}
to be solved in $\theta$.

\paragraph{Solving $A\theta=b$ where $A$ and $b$ are samplable.} An
obvious way to solve $A\theta=b$ via a stochastic gradient method is
\begin{equation}
\theta\gets \theta - \eta (A\theta-b)
\end{equation}
using samples for $A$ and $b$: the only fixed point solves $A\theta=b$.
TD does exactly this, 
with $A=\transp{\Phi}\rho(\Id-P)\Phi$ and $b=\transp{\Phi}\rho R$.

If $A$ is symmetric positive definite, this is also the
gradient descent of the loss function
$\transp{(A\theta-b)}A^{-1}(A\theta-b)$. However, we want to apply this
to matrices $A$ that are not necessarily of this type.

Under this update for $\theta$, the error $\theta-A^{-1}b$ is updated as
\begin{equation}
\theta-A^{-1}b \gets (\Id-\eta A)(\theta-A^{-1}b)
\end{equation}

This converges if only if $A$ is \emph{stable}.

\begin{defi}
A matrix $A$ is \emph{stable} if one of the following equivalent
conditions are satisfied:
\begin{itemize}
\item All the eigenvalues of $A$ have positive real part.
\item The matrix $(\Id-\eta A)$ has spectral radius $<1$ for small enough
$\eta$.
\item For $\eta$ small enough, iterating $\theta\gets (\Id-\eta A)\theta$
converges to $0$.
\item The differential equation $\theta'=-A\theta$ converges to $0$ for
any initial value.
\item There exists a symmetric, positive definite matrix $L$ (Lyapunov
function) such that $\transp{\theta}L\theta$ is decreasing along the
solutions of the differential equation $\theta'=-A\theta$.
\item There exists a symmetric, positive definite matrix $L$ such that
\begin{equation}
\transp{\theta} L A \theta>0
\end{equation}
for all $\theta\neq 0$. (This is the same $L$ as in the previous
condition.)
\end{itemize}
\end{defi}

In particular, a symmetric, definite positive matrix is stable (last
criterion with $L=\Id$).
Stability is invariant by matrix similarity $A\gets B^{-1}A B$, since
this preserves eigenvalues.

Since the
solution of $\theta_t'=-A\theta_t$ is $\theta_t=e^{-tA}\theta_0$, a
Lyapunov function that works is $\transp{\theta_0}L\theta_0\deq \int_{t \geq 0}
\norm{\theta_t}^2$. Indeed this is decreasing, because
$\transp{\theta_t}L\theta_t$ is just the same integral starting at $t$
instead of $0$. Explicitly this is
$L=\int_{t\geq 0} \transp{(e^{-t A})}
e^{-tA}$.

\paragraph{Making the TD iteration stable.} In linear TD the matrix $A$ is
\begin{equation}
A\deq \transp{\Phi}\rho (\Id-P)\Phi
\end{equation}
and we want to solve
\begin{equation}
A\theta=b\,\qquad b\deq \transp{\Phi}\rho R
\end{equation}
If $A$ is stable, then iterating $\theta\gets \theta -\eta(A\theta-b)$ will
converge.

We know $A$ is stable in several cases: (TODO here I just ignore constant
functions)
\begin{enumerate}
\item On-policy: $\rho=\mu$, the invariant distribution. Then we know that $\rho(\Id-P)$ is positive
(Dirichlet form: $\transp{f}\rho(\Id-P)f=\E(f(s_t)-f(s_{t+1}))^2$). This implies that $A=\transp{\Phi}\rho (\Id-P)\Phi$ is
positive too. So we can take $L=\Id$ and we recover convergence of linear
TD on-policy.
\item In the tabular case: $\Phi=\Id$. Indeed we have $A=\rho (\Id-P)$,
which is not positive in general. But $\mu (\Id-P)$ is positive, so we can take
$L=\mu/\rho$ so that $LA$ is positive. Thus, tabular TD works
off-policy.
\end{enumerate}

One can also \emph{precondition} the gradient descent. This can make it
stable and/or make convergence faster. The step becomes
\begin{equation}
\theta\gets \theta-\eta B(A\theta-b)
\end{equation}
for some matrix $B$. Some choices of $B$ correspond to natural choices of
loss functions on $\theta$, as we will see.

Of course the values of $B$ that are most interesting are obtained as the
inverse of something; for instance $B=A^{-1}$ solves the problem
instantly... Two-timescale methods or TANGO-like algorithms
\footnote{TANGO looks much like a two-timescale algorithm, but is
``almost one-timescale'' in that the faster timescale can be fixed and
does not need to tend to $0$.} can apply an
inverse of something to $A\theta-b$, via an auxiliary gradient descent.
TANGO works for stable matrices, if they are samplable \footnote{In
TANGO, the consecutive steps absolutely have to be \emph{independent}. So even in
the on-policy case, we will need a replay buffer from $\mu$ for the auxiliary
gradient descent.}

A few choices are:
\begin{itemize}
\item GTD is a gradient descent of
$\transp{(A\theta-b)}(A\theta-b)$ (Sutton 2009). This helps with $A$ being non-symmetric.
This corresponds to $B=\transp{A}$. Then
$BA=\transp{A}A$ is
obviously stable, hence convergence.

TODO: check if
GTD uses the same sample for $\transp{A}$ and $A$. If so, it will
converge to a different stable point. TODO: Can be applied TANGO-style
(by formally inverting $\Id$ in TANGO), advantage=fewer variance due to
two-sampling (Is this equivalent to eligibility traces without the
variance??). (TANGO uses two-sampling but on a sum of past terms, so
that a past term gets TANGOified whenever we happen to reach a state that
activates the same features in the future, which is probably fine). Can
also be TANGOified with $\gamma\to 0$ (two-timescale) which now requires
only single sampling (no replay buffers).

\item $B=A^{-1}$: obviously stable ($BA=\Id$), jumps to the solution in one step if
$\eta=1$. Requires computing the inverse of $A$, namely, solving the
problem by first solving the problem. But if $A$ is stable and samplable,
this can be done via TANGO. More on that later. This should speed up
things in situations where $A$ is stable in the first place (eg.,
on-policy).

\item $B=(\transp{\Phi}\rho\Phi)^{-1}$. This amounts to orthonormalizing
the basis before applying TD. In general, this is not stable off-policy: take the
standard two-point counterexample with a single $\Phi=\transp{(1,2)}$,
orthonormalizing $\Phi$ is not going to help as $\transp{\Phi}\rho\Phi$
is just a number.

Still, this is stable in two cases:
\emph{on-policy} (take $L=B^{-1}$, given that $A$ is stable on-policy), or
assuming \emph{overparameterization} ($\Phi$ invertible: then we can take
$L=\transp{\Phi}\mu \Phi$ so that $LBA=\transp{\Phi}\mu (\Id-P)\Phi$
which is positive because $\mu(\Id-P)$ is positive; essentially, if
$\Phi$ is invertible we are back to the tabular case).
% Then we have to prove th
% $(\transp{\Phi}\Phi)^{-1}\transp{\Phi}\rho(\Id-P)\Phi$ is stable. Taking
% $L=\transp{\Phi}(\mu/\rho)\Phi$, and computing
% \begin{align}
% \transp{\theta}LBA\theta&=
% \transp{\theta}L(\transp{\Phi}\rho\Phi)^{-1}\transp{\Phi}\rho(\Id-P)\Phi\theta
% \end{align}
% and note that $(\transp{\Phi}\rho\Phi)^{-1}\transp{\Phi}\rho$ is the
% solution of linear regression with design matrix $\Phi$ in $L^2(\rho)$.
% This implies that
% \begin{equation}
% \Phi(\transp{\Phi}\rho\Phi)^{-1}\transp{\Phi}\rho=\Pi_\Phi
% \end{equation}
% where $\Pi$ is the orthogonal projector on the image of $\Phi$ in
% $L^2(\rho)$. Now taking
% \begin{equation}
% L=\transp{\Phi}(\mu/\rho)\Phi
% \end{equation}
% we have
% \begin{equation}
% \transp{\theta}LBA\theta=\transp{\theta} \transp{\Phi}(\mu/\rho)\Pi_\Phi
% (\Id-P)\Phi\theta
% \end{equation}
% TODO: on-policy or overparameterized

\item The choice from Sutton 2009 "TDC" combines the previous two tricks
and corresponds to a gradient descent
of the projected norm of the Bellman gap on the subspace generated by
$\Phi$, projected in $L^2(\rho)$. This is GTD with $L^2(\rho)$-orthonormalization of
$\Phi$, namely,
\begin{equation}
B=\transp{A} (\transp{\Phi}\rho\Phi)^{-1}=\transp{(\transp{\Phi}\rho(\Id-P)\Phi)}(\transp{\Phi}\rho\Phi)^{-1}
\end{equation}
from which it is clear that $BA$ is symmetric and stable. This is obtained as a
gradient descent of the projected norm of the Bellman gap on the image of
$\Phi$, projected in $L^2(\rho)$. TODO: TDC and GTD2 differ by using that
$A$ has $(\Id-P)$ on  the right, and expanding the $\Id$ term and the $P$
term separately, which results in simplifications and make TDC much
closer to TD and less reliant on the inversion.

\item The choice $B=A^{-1}$ can also be written as
$B=\transp{A}(A\transp{A})^{-1}$. The advantage is that we now have to
invert $(A\transp{A})$ which is a stable matrix, so, better behaved with
TANGO or two-timescale. However, variance
will be terrible: this requires four independent samples from $A$.

\item A gradient descent of
$\transp{(A\theta-b)}(A+\transp{A})^{-1}(A\theta-b)$, corresponding to
$B=\transp{A}(A+\transp{A})^{-1}$. This is probably the closest to
unmodified TD in behavior: for instance when $A$ is symmetric ($\rho=\mu$
and $P$ reversible) this coincides with TD. $A+\transp{A}$ corresponds to
the Dirichlet form for $\rho$, namely, $\E_{\rho} (f(s)-f(s'))^2$ TODO
CHECK.

\end{itemize}

Only $B=A^{-1}$ and $B=\transp{A}(A\transp{A})^{-1}$ are truly
second-order (namely, solve the problem in one step in the non-stochastic
case).

The other algorithms involving an inverse basically have the cost of
second-order algorithms without the full benefits.  They are second-order
in feature-space (they orthonormalize $\Phi$) but not in Markov process
space (they don't backpropagate rewards).

TODO sampling: if $A=\E \tilde A$, can do $\transp{(\E \tilde A)}(\E
\tilde A)$ or
$\E(\transp{\tilde A}\tilde A)$, which is not equal to $\transp{A}A$ but
is still definite positive (hence stable). This would avoid some
quadruple-sampling problems with $B=\transp{A}(A\transp{A})^{-1}$. But
breaks second-order correctness?

TODO: define second-order with respect to $\Phi$ (second-order
provided $\transp{\Delta}\Delta=\Id$, eg in a Fourier frequency-adjusted
basis), with respect to the Markov chain (second-order provided
$\transp{\Phi}\Phi=\Id$).

Almost-one-timescale (such as TANGO) needs more independence than two-timescale. I
\emph{think} that two-timescale can deal with taking the same sample for
$A$ and $\transp{A}$ in $\transp{A}(...)^{-1}A$, and taking correlated
consecutive samples (eg, online) for the matrix to be inverted. I think
TANGO would need uncorrelated samples for the matrix to be inverted (eg,
a replay buffer) and two samples for $A$ and $\transp{A}$, but not sure
about the latter.

TODO: I'm more interested in an algo that's second-order wrt the Markov
process (I don't necessarily want to orthonormalize features).

TODO: in the original TANGO paper: prove that the main learning rate
doesn't have to be too small even for non-quadratic problems, namely,
even if the Hessian changes, we're still applying a positive operator to
earlier gradients so we still have a valid gradient direction.

\paragraph{Nonlinear case.} In the nonlinear case, we still set $\Phi\deq
\partial V/\partial \theta$ and TD is
\begin{equation}
\theta \gets \theta-\eta \transp{\Phi}\rho (V-PV-R)
\end{equation}
and the resulting change of $V$ is
\begin{equation}
V\gets V-\eta \Phi\transp{\Phi}\rho (V-PV-R)+O(\eta^2)
\end{equation}

Preconditioning TD with a matrix $B$ amounts to using $B\transp{\Phi}$ instead of
$\transp{\Phi}$ in the update of $\theta$, and $\Phi B \transp{\Phi}$ instead of
$\Phi\transp{\Phi}$ in the update of $V$.

Let us consider a loss function $\transp{(V-PV-R)}L(V-PV-R)$ on the
Bellman gap, with $L$ a symmetric definite positive matrix. For small
$\eta$, the derivative of this loss function wrt $\theta$ is
\begin{equation}
\transp{(V-PV-R)}L(\Phi-P\Phi)
\end{equation}

The loss function $L$ may not depend on $\theta$, otherwise we are
decreasing a variable Lyapunov function which does not prove convergence.

If the algorithm is some form of preconditioned TD, namely
$\delta\theta=-\eta B\transp{\Phi}\rho (V-PV-R)$, the loss function will
decrease if and only if
\begin{equation}
\transp{(V-PV-R)}L(\Phi-P\Phi)B\transp{\Phi}\rho (V-PV-R)
\end{equation}
is positive, namely, if $L\Delta \Phi B \transp{\Phi}\rho $ is a positive
operator.

On-policy, a choice would seem to be $L=\mu\Phi B \transp{\Phi}\mu$ which
would result in $(\mu\Phi B \transp{\Phi})(\mu \Delta) \mu\Phi B
\transp{\Phi}\mu$ which is positive; however, this depends on $\theta$
since $\Phi$ does. So this is not a suitable Lyapunov function.

In the reversible case, $\mu\Delta$ is symmetric.\footnote{$\rho \Delta$
can be symmetric if and only if $\rho=\mu$. Indeed, if $\rho \Delta$ is
symmetric then $\rho=\rho'$ so that $\rho$ is the stationary distribution.} We can use the
Dirichlet form which amounts to $L\deq\mu(\mu\Delta)^{-1}\mu$; then
$L\Delta=\mu$ and we have to prove that $\mu\Phi B\transp{\Phi}\rho$ is
positive.  On-policy, $\rho=\mu$ and this is obviously true, for any
positive $B$. Thus, in the reversible case, nonlinear TD converges
on-policy, for any $B$. It does so by minimizing the Dirichlet norm.

In the non-reversible case, let $\rho\Delta=S+A$ with $S$ symmetric and
$A$ antisymmetric. Assume $S$ is positive. Let $L\deq \rho S^{-1} \rho$. Then we have to prove that
$\rho S^{-1} \rho \Delta \Phi B\transp{\Phi}\rho$ is positive. For any
function $f$,
\begin{align}
\transp{f} \rho S^{-1} \rho \Delta \Phi B\transp{\Phi}\rho f
&=\transp{f} \rho S^{-1} (S+A) \Phi B\transp{\Phi}\rho f
\\&=\transp{f} \rho \Phi B\transp{\Phi}\rho f + \transp{f} \rho S^{-1}
A \Phi B\transp{\Phi}\rho f
\end{align}

The first term $\transp{f}
\rho \Phi B\transp{\Phi}\rho f$ is always positive. For the second
term, since $A$ is antisymmetric, we have $\transp{x}Ax=0$ for all $x$.
Thus, we can add any multiple of $\transp{x}Ax=0$; let us take
$x=S^{-1}\rho f$:
\begin{equation}
\transp{f} \rho S^{-1}
A \Phi B\transp{\Phi}\rho f=
\transp{f} \rho S^{-1}A (\Phi B\transp{\Phi}-S^{-1})\rho f
\end{equation}

In particular, if $S^{-1}=\Phi B\transp{\Phi}$ then this vanishes.
The latter can only happen in the overparameterized case (indeed, the
right-hand-side lives in the image of $\Phi$). In that case $\Phi$ is
invertible and $B=(\transp{\Phi}S\Phi)^{-1}$ solves the problem; a
two-timescale algorithm is required to compute $B$.

Thus, assuming $S$ is positive, we have a two-timescale algorithm for non-linear TD in the off-policy
overparameterized case; it minimizes the Dirichlet norm.
However, in general $S$ is not positive.

Still, $S$ is positive if $\rho=\mu$, so,
we have a two-timescale algorithm for \emph{non-linear on-policy
overparameterized TD}.

(Still works for non-overparameterized if the antisymmetric part $A$ does not couple the image
of $\Phi$ and its complement...)

TODO: Why do I end up using $\rho\Delta+\transp{(\rho \Delta)}$ while the
Dirichlet form is $\rho \Delta + \rho'-\transp{(\rho P)}$ which is
different, I think...  Probably use the right combination of $\rho$ and
$\rho'$ and the Dirichlet form rather than $\rho S^{-1}\rho$.

Nonlinear GTD2 etc use a function $L$ that depends on $\theta$ (via
$\Phi$); this adds Hessian terms to the gradient of $L$, and it's not
obvious how to interpret the local minima of $L$.

\paragraph{Why $S$ is not positive.} The Dirichlet form associated
with the set of transitions $s\to s'$ is
\begin{align}
\Dir&=\rho+\rho'-\rho P-\transp{(\rho P)}
\\&=\E[\1_{ss}+\1_{s's'}-\1_{ss'}-\1_{s's}]
\end{align}
where the expectation is over transitions in the dataset. The Dirichlet
form satisfies
\begin{equation}
\qquad \transp{f}\Dir f=\E_{ss'}(f(s)-f(s'))^2
\end{equation}
which is obviously positive. Meanwhile,
\begin{align}
\rho\Delta&=\rho-\rho P=\E[\1_{ss}-\1_{ss'}]
\\&=\frac12 \Dir+\frac12 \E[\1_{ss}-\1_{s's'}]+\frac12
\E[\1_{s's}-\1_{ss'}]
\end{align}
and the first two terms are symmetric while the last term is
antisymmetric, so
\begin{equation}
S=\frac12 \Dir+\frac12 \E[\1_{ss}-\1_{s's'}],\qquad A=\frac12
\E[\1_{s's}-\1_{ss'}]
\end{equation}
and
\begin{equation}
\transp{f}Sf=\E f(s)^2-\E f(s)f(s')
\end{equation}
which in general is not positive (take $\rho$ concentrated on a single $s$, and a large value
for $f(s')$).

\paragraph{Minimizers of errors.} Rather than the fixed point of TD, one
can look for minimizers of the error in various norms.

A local minimizer of the $L^2(\rho)$ error between $V$ and
$(\Id-P)^{-1}R$ solves
\begin{equation}
\transp{\Phi} \rho V=\transp{\Phi}\rho (\Id-P)^{-1} R
\end{equation}
which is not quite the same as the fixed point of TD. We have no access
to the right-hand-side, indeed, computing $(\Id-P)^{-1} R$ is exactly the
original problem.

A local minimizer of the
$L^2(\rho)$ error between $(\Id-P)V$ and $R$ is
\begin{equation}
\transp{\Phi}\transp{(\Id-P)}\rho(\Id-P)V=\transp{\Phi}\transp{(\Id-P)}\rho R
\end{equation}
and in the linear case $V=\Phi\theta$, this is the standard solution to
the linear regression problem $(\Id-P)\Phi\theta\approx R$ in
$L^2(\rho)$. However, the matrix on the left is not samplable: sampling
$\transp{(\Id-P)}\rho(\Id-P)$ is exactly equivalent to double sampling
from $s\sim \rho$.
Indeed, $\transp{P}\rho P$ is equal to $\E[\1_{s' s''}]$ where $s'$ and
$s''$ are two independent states from a transition $s\to s'$ with $s\sim
\rho$.

The minimizer of the $L^2(\rho^2)$ error between $R$
and $(\Id-P)V$ is samplable:
\begin{equation}
\transp{\Phi}\transp{(\Id-P)}\rho^2(\Id-P)V=\transp{\Phi}\transp{(\Id-P)}\rho^2
R
\end{equation}
Indeed, denoting $\Delta\deq \Id-P$, we can sample $\rho \Delta$ and
therefore we can take two samples from $\rho \Delta$ which provides a
sample from $\transp{(\rho\Delta)}\rho\Delta$. However, the double sampling
problem resurfaces: Most of the time, taking the product
$\transp{(\rho\Delta)}\rho\Delta$ from two samples of $\rho
\Delta$ is $0$ unless we are lucky enough to sample two transitions $s\to
s'$ with the same $s$.
Even the right-hand-side $\rho^2 R$ involves double sampling.

If one has access to some features $\Psi$, this can be mitigated:
the minimizer of the norm $L^2(\rho \Psi\transp{\Psi}\rho)$ is better
samplable, because now in double sampling it's enough to sample states \emph{which share
some
features}, not just the exact same state. This is basically what happens
in the linear case, with $\Psi=\Phi$. In the nonlinear case taking
$\Psi=\Phi$ is not acceptable because $\Phi$ depends on $\theta$, so the
objective is not a fixed norm to be minimized. The features $\Psi$ should not be orthogonal. Should use a
hash of states instead? Or apply ReLU to orthogonal features.

The Dirichlet norm associated with $\rho$ is $\E_{ss'}
(f(s)-f(s'))^2$. Its matrix is $\rho+\rho'-\rho P-\transp{(\rho P)}$ where
$\rho'$ is the distribution of $s'$ (and is samplable).
A local minimizer of the Dirichlet error between $V$ and
$(\Id-P)^{-1}R$ solves
\begin{equation}
\transp{\Phi} (\rho+\rho'-\rho P-\transp{(\rho P)})(V-(\Id-P)^{-1}R)=0
\end{equation}
If $\rho$ is the stationary distribution $\mu$ and $P$ is reversible, one
has $\transp{(\mu P)}=\mu P$ and $\rho'=\mu$ so $\rho+\rho'-\rho P-\transp{(\rho
P)}=2\mu(\Id-P)$ and this is equivalent to
\begin{equation}
\transp{\Phi}\mu ((\Id-P)V-R)=0
\end{equation}
namely, the same as a fixed point of TD.

Instead of applying the Dirichlet norm to $(V-(\Id-P)^{-1}R)$, one can
apply the inverse Dirichlet norm to $((\Id-P)V-R)$. When $P$ is
reversible and $\rho=\mu$ this
is the same. One tries to minimize
$\transp{((\Id-P)V-R)}\rho\Dir^{-1}\rho((\Id-P)V-R)$ where
$\Dir\deq \rho+\rho'-\rho P-\transp{(\rho P)}$ as above, and use
TANGO to invert $\Dir$. However this suffers again from a double
sampling problem, in several places. Still, this can be rewritten as a
correction to TD where only the correction suffers from double
sampling/large variance. Namely, the gradient of the above with respect
to $\theta$ is $\transp{\Phi}\transp{(\rho\Delta)} \Dir^{-1}\rho(\Delta V-R)$. Since
$\rho\Delta$ is half of $\Dir$, this gradient is 
\begin{gather}
\transp{\Phi}\transp{(\rho\Delta)} \Dir^{-1}\rho(\Delta V-R)=
\\
\frac12\transp{\Phi}(\transp{(\rho \Delta)}+\rho'-\rho P)\Dir^{-1}\rho(\Delta
V-R)+\frac12 \transp{\Phi}(\transp{(\rho \Delta)}+\rho P-\rho')\Dir^{-1}\rho(\Delta
V-R)
\\=\frac12\transp{\Phi}\rho(\Delta
V-R)+\frac12 \transp{\Phi}(\transp{(\rho \Delta)}+\rho P-\rho')\Dir^{-1}\rho(\Delta
V-R)
\end{gather}
the first part of which is TD. Thus one can write a correction to TD such
that only the correction suffers from double-sampling. The correction vanishes for reversible
chains. The double sampling can be simulated with high variance in finite
state spaces (just take samples from $P$ and $\transp{P}$ until they
match) but does not make sense in continuous state spaces. Use a model of
transitions only for the correction? TODO: use the same trick on the
right? No, because the $R$ term doesn't go away and I'd rather have a
correction that vanishes when $\Delta V-R$ is small.

\section{A function that almost solves the Bellman equation can be very
far from the true Bellman function}

Combined with all these problems, is the problem that a Bellman solution on the
support of $\rho$ may be very, very far from the global Bellman solution,
even on the support of $\rho$ (take random walk on a circle, with $\rho$
being everything but one point: then any linear function solves Bellman
on the support of $\rho$).

TODO seems to be much worse off-policy.

\section{Storing a summary of possible transitions?}

The derivative of the average reward with respect to the policy
parameters $\psi$ is equal to
\begin{align}
\partial_\psi \left(\E_\mu r\right)&=
\E_{s,a,s'} \partial_\psi \log \pi(a|s)\left(
r+V(s')
\right)
\\&=\E_{s,a}\partial_\psi \log \pi(a|s)\E_{s'|s,a}\sum_{k\geq 0}P^k R(s')
\\&=\E_{s,a,s'}\partial_\psi \log \pi(a|s)(\Delta^{-1}R)(s')
\end{align}
by expanding $V(s')$ as a sum of expected future rewards (and attributing
the reward of a transition $s\to s'$ to $s'$, which makes sense if $s'$
is a sufficient statistic).
(Here I'm neglecting constant functions again.)

Thus if we define the vector
\begin{equation}
g\deq \E_{s,a,s'}\partial_\psi \log \pi(a|s)\1_{s'}
\end{equation}
we have
\begin{equation}
\partial_\psi \left(\E_\mu r\right)=\transp{g}\Delta^{-1}R
\end{equation}
Let us define the row vector
\begin{equation}
G\deq \transp{g}\Delta^{-1},\qquad \partial_\psi \left(\E_\mu
r\right)=GR
\end{equation}
then one has
\begin{equation}
G(s)=\partial_\psi \mu(s)
\end{equation}
namely, $G$ encodes the derivative of the invariant distribution with
respect to the policy parameters. Thus, if we find a state $s$ with high
reward $R(s)$, $G$ tells us how we can visit $s$ more often. This
explains that the derivative of the average reward is $GR$.

[$g$ sums to $0$ on-policy, so we don't have to worry about removing
constants and can directly take $\gamma=1$: $\transp{g}\Delta^{-1}$ is
well-defined.]

Eligibility traces are a random, unbiased estimator of $G$: if
$(s_t)_{t\in \Z}$ is a
trajectory from the current policy then
\begin{equation}
\E\sum_{k\geq 1}\partial_\psi \log \pi(a_{t-k}|s_{t-k})\1_{s_t}=G
\end{equation}
Moreover $G$ satisfies a backward Bellman equation on-policy,
\begin{equation}
G=\E_{s,a,s'} \partial_\psi \log \pi(a|s)\1_{s'}+GP
\end{equation}
thanks to $\Delta^{-1}=\sum P^k=\Id+\Delta^{-1}P$. Eligibility traces are
precisely
updated this way at each step, via a sampling of $P$.

$G$ cannot be stored but one can store an approximate model. For instance
one could approximate $G$ as a sum of parameters times functions:
\begin{equation}
G\approx \sum_i \psi_i \phi_{\theta_i}(s)
\end{equation}
where each $\psi_i$ is a policy parameter, and each $\phi_{\theta_i}$ is
a function of the state parameterized by some model. Then the reverse
Bellman equation above allows us to learn $\psi_i$ and $\theta_i$ via 
a TD-like algorithm, \emph{on-policy}.

TODO: not the right representation, plus works only on-policy. Maybe learn a function $a(s,s')$ that
returns the first action to be taken if we want to reach $s'$ from $s$?

\paragraph{Learning to navigate.} Define a function $V(s,s')$ that
represents the time spent in $s'$ if starting at $s$ and acting optimally
towards $s'$ (just as if the reward was $\1_{s'}$). It is defined by an iterated expectimax to reach $s'$, and
satisfies the Bellman equation
\begin{equation}
V(s,s')=\1_{s=s'}+\gamma \max_a \E_{s_1|s,a} V(s_1,s')
\end{equation}
and it also satisfies
\begin{equation}
V(s,s')=\1_{s=s'}+\gamma \sum_{s_{t-1}} V(s,s_{t-1})\max_a
\E_{s_t|s_{t-1},a}\1_{s_t=s'}
\end{equation}
although it's not clear to me how to exploit the last expression. 
\footnote{On-policy it works, I think, because on-policy we can sample
reverse transitions. But I think in that case $V(s,s')$ has to be defined
as a density wrt $\mu(s')$.\\Should also work if we maintain a low-rank
estimate of $V$ as a sum of Diracs over a few $s'$.}
But I
think we have to work from both ends if we want to represent
compositionality of actions. Indeed with the first equation, different values
of $s'$ are just completely separate Bellman problems; the generalization
abilities of the function approximator for $V$ may help, but only so
much. Each time we observe a transition $s\to s'$, we should be able to
update the probabilities to reach $s'$ based on the probabilities to
reach $s$.

Then we may learn a goal-oriented policy $\pi_\theta(a|s,s')$ that returns an
action $a$ at state $s$ which produces a move towards $s'$. Learning
$\pi_\theta$ may be done by observing transitions and maximizing
$V(s,s')$.

It's not clear either how to exploit $V(s,s')$ to maximize a reward. If
we were considering a fixed policy we would have
$V(s)=\sum_{s'}V(s,s')R(s')$ (namely $V(s,s')$ would be the value
function of the reward function $\1_{s'}$). But here $V(s,s')$ uses a
strategy that depends on the goal $s'$.

Another possibility is to learn a goal-oriented policy towards desired
\emph{features} of the final state $s'$. Given a basis $(\Psi_i)$ of
features, and a set of coefficient $\alpha_i$, one might want
$\pi_\theta(a|s,(\alpha_i))$ to produce a policy that maximizes the
expected reward $\sum \alpha_i \Psi_i$ (thus, simultaneously learning to
solve a number of RL problems). 
This can be done either via learning a value function $V(s,(\alpha_i))$
then actor-critic, or via an eligibility-trace-like thing (but then it
has to be on-policy, for which policy??)

Then we can learn an approximate decomposition of the actual reward over
the $\Psi_i$, and we know what to do. (This assumes the reward is well
approximated by the $\Psi_i$; or that we learn the complement.) Advantage
= even if there is no reward signal for some time, we still learn to
navigate the environment.

There are many ways to build interesting
$\Psi_i$'s: hidden features of a world model/reward model, eigenfunctions
of $P$, or just setting a basis of random time-dependent functions on a
trajectory and smoothing them via a NN (``zeitgeist'').

(The relationship between $V(s,s')$ and $V(s,(\alpha_i))$ is: the latter
is more general, by taking a full basis of Dirac features $\1_{s'}$. Then
$V(s,(\alpha_i))$ does not just learn how to go to individual states, but
how to maximize time spent in an arbitrary weighted sum of states. This
is why we can later extend to an arbitrary reward function.)

See \emph{Universal value function approximators},
Schaul--Horgan--Gregor--Silver ICML 2015

\end{document}