### 2019-01-12 paper/icml_drau/pendulum_fig.tex
First figure

%! TEX root = icml_drau.tex
\newcommand{\iw}{1cm}
\begin{figure*}[!ht]
	\begin{tabular}{c|c}
		DDPG & AU \\
		\hline
		\begin{tabular}{l|lllll}
			$\delta t = 0.01$ & \includegraphics[width=\iw]{figs_data/pendulum/0.01/ddpg_act_2.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/ddpg_act_4.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/ddpg_act_6.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/ddpg_act_8.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/ddpg_act_10.png} \\
			$\delta t = 0.005$ &\includegraphics[width=\iw]{figs_data/pendulum/0.005/ddpg_act_4.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/ddpg_act_8.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/ddpg_act_12.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/ddpg_act_16.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/ddpg_act_20.png} \\
			$\delta t = 0.002$ &\includegraphics[width=\iw]{figs_data/pendulum/0.002/ddpg_act_10.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/ddpg_act_20.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/ddpg_act_30.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/ddpg_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/ddpg_act_50.png} \\
			$\delta t = 0.001$ &\includegraphics[width=\iw]{figs_data/pendulum/0.001/ddpg_act_20.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/ddpg_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/ddpg_act_60.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/ddpg_act_80.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/ddpg_act_100.png} \\
			$\delta t = 0.0005$ &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/ddpg_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/ddpg_act_80.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/ddpg_act_120.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/ddpg_act_160.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/ddpg_act_200.png}
		\end{tabular} &
		\begin{tabular}{lllll}
			 \includegraphics[width=\iw]{figs_data/pendulum/0.01/au_act_2.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/au_act_4.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/au_act_6.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/au_act_8.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/au_act_10.png} \\
			\includegraphics[width=\iw]{figs_data/pendulum/0.005/au_act_4.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/au_act_8.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/au_act_12.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/au_act_16.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/au_act_20.png} \\
			\includegraphics[width=\iw]{figs_data/pendulum/0.002/au_act_10.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/au_act_20.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/au_act_30.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/au_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/au_act_50.png} \\
			\includegraphics[width=\iw]{figs_data/pendulum/0.001/au_act_20.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/au_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/au_act_60.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/au_act_80.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/au_act_100.png} \\
			\includegraphics[width=\iw]{figs_data/pendulum/0.0005/au_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/au_act_80.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/au_act_120.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/au_act_160.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/au_act_200.png}
		\end{tabular}
	\end{tabular}
	\caption{Policies obtained by DDPG and AU at different macroscopic instants of training. Each image represents the policy learnt by the policy network, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the closer to $1$ the action, the darker, the closer to $-1$.}
\end{figure*}