### 2019-01-18 paper/icml_drau/pendulum_fig.tex
Commiting figures.

%! TEX root = icml_drau.tex
\newcommand{\iw}{1cm}
\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{figs_data/pendulum_fig.pdf}
	\caption{Policies obtained by DDPG and AU at different macroscopic instants of training. Each image represents the policy learnt by the policy network, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the closer to $1$ the action, the darker, the closer to $-1$.}
\end{figure*}