### 2019-01-19 paper/icml_drau/pendulum_fig.tex

%! TEX root = icml_drau.tex
\newcommand{\iw}{1cm}
\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{figs_data/pendulum_fig.pdf}
	\caption{Policies obtained by DDPG and AU at different instants in physical time (or normalized epochs) of training on the pendulum swing-up environment. Each image represents the policy learnt by the policy network, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the closer to $1$ the action, the darker, the closer to $-1$.}
        \label{fig:pend}
\end{figure*}