### 2019-01-21 paper/icml_drau/supplementary.tex

\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{xcolor}

\newtheorem{theorem}{Theorem}
\input{macros.tex}

\begin{document}
\section{Proofs}
We hereby give proofs for all the results presented in the paper.
The first result presented is a proof of convergence for discretized
trajectories.
\begin{theorem}
	Let $F: {\cal S} \times {\cal A} \rightarrow \bb{R}^n$ and $\pi: {\cal S}
	\rightarrow {\cal A}$ be the dynamic and policy functions. Assume that,
	for any $a$, $s \rightarrow F(s, a)$ and $s \rightarrow F(s, \pi(s))$
	are ${\cal C}^1$, bounded and $K$-lipschitz.  For
	a given $s_0$, define the trajectory $(s_t)_{t\geq 0}$ as the unique
	solution of the differential equation
	\begin{equation}
		\frac{ds_t}{dt} = F(s_t, \pi(s_t)).
		\label{eq:diff}
	\end{equation}
	For any $\deltat > 0$, define the discretized trajectory
	$(s_\deltat^k)_k$ recurrently as $s_\deltat^0 = s_0$,
	$s_\deltat^{k + 1}$ is the value at time $\deltat$ of
	the unique solution of
	\begin{equation}
		\frac{d\tilde{s}_t}{dt} = F(\tilde{s}_t, \pi(s_\deltat^k))
	\end{equation}
	with initial point $s_\deltat^k$.
	Then, for every $t \geq 0$
	\begin{equation}
		\|s_t - s_\deltat^{\lfloor \frac{t}{\deltat} \rfloor}\|
		\leq 
	\end{equation}
	\label{th:traj-conv}
\end{theorem}
\begin{proof}
	\input{proof_traj.tex}
\end{proof}
\begin{theorem}
	Under suitable assumptions \TODO{ref appendix}, for all $s \in {\cal
	S}$, one has
	%\begin{equation}
	%\label{eq:conv-value}
	$V^\pi_\deltat(s) = V^\pi(s) + \bigO(\deltat)$
	%\end{equation}
	when $\deltat\to 0$.
	\label{th:conv-value}
\end{theorem}
\begin{proof}
	\input{proof_conv_value.tex}
\end{proof}
\begin{theorem}
	$A^\pi_\deltat$ convergence theorem.
\end{theorem}
\begin{proof}
	\input{proof_conv_adv.tex}
\end{proof}
\begin{theorem}
	Discrete parameter trajectories converge to continuous trajectories.
\end{theorem}
\begin{proof}
	\input{proof_conv_params.tex}
\end{proof}
\section{Implementation details}
\end{document}