### 2019-01-21 paper/icml_drau/cdau.tex
.

%! TEX root = supplementary.tex

\begin{algorithmic}
	\STATE \textbf{Inputs:}
	\STATE $\theta$, $\psi$ and $\phi$, parameters of
	$V_{\theta}$, $\bar{A}_{\psi}$ and $\pi_\phi$.
	\STATE $\pi^{\text{explore}}$ and $\nu_\deltat$ defining an exploration policy.
	\STATE \textbf{opt}$_V$, \textbf{opt}$_A$, \textbf{opt}$_\pi$, $\alpha^V \deltat$, $\alpha^A \deltat$ and $\alpha^\pi\deltat$, optimizers and learning rates.
	\STATE $\mathcal{D}$, buffer of transitions $(s, a, r, d, s')$, with $d$ the episode termination signal.
	\STATE $\deltat$ and $\gamma$, time discretization and discount factor.
	\STATE \textbf{nb\_epochs} number of epochs.
	\STATE \textbf{nb\_steps}, number of steps per epoch.
	\STATE
	\STATE Observe initial state $s^0$
	\STATE $t \gets 0$
	\FOR {$e=0, \textbf{nb\_epochs}$}
	\FOR {$j=1, \textbf{nb\_steps}$}
	\STATE $a^k \leftarrow \pi^{\text{explore}}(s^k, \nu^k_\deltat)$.
	\STATE Perform $a^k$ and observe $(r^{k+1}, d^{k+1}, s^{k+1})$.
	\STATE Store $(s^k, a^k, r^{k+1}, d^{k+1}, s^{k+1})$ in $\mathcal{D}$.
	\STATE $k \gets k + 1$
	\ENDFOR
	\FOR {$k=0, \text{nb\_learn}$}
	\STATE \text{Sample a batch of $N$ random transitions from $\mathcal{D}$}
	\STATE $Q^i \gets V_{\theta}(s^i) + \deltat\hspace{-.17em}\left(
	\bar{A}_{\psi}(s^i, a^i) - \bar{A}_{\psi}(s^i, \pi_\phi(s^i))\right)$
	\STATE $\tilde{Q^i} \gets r^i\deltat + (1 - d^i) \gamma^{\deltat} V_{\theta}(s'^i)$
	\STATE $\Delta \theta \gets \frac{1}{N}\sum\limits_{i=1}^N  \frac{\left(Q^i - \tilde{Q^i}\right)\partial_{\theta} V_{\theta}(s^i)}{\deltat}$
	\STATE $\Delta \psi \gets \frac{1}{N}\sum\limits_{i=1}^N \frac{\left(Q^i - \tilde{Q^i}\right)\partial_{\psi} \left(\bar{A}_{\psi}(s^i, a^i) - \bar{A}_{\psi}(s^i, \pi_\phi(s^i))\right) }{\deltat}$
	\STATE $\Delta \phi \gets \frac{1}{N} \sum\limits_{i=1}^N \partial_a \bar{A}_\psi(s^i, \pi_\phi(s^i)) \partial_\phi \pi_\phi(s^i)$
	\STATE Update $\theta$ with \textbf{opt}$_V$, $\Delta \theta$ and learning rate $\alpha^V \deltat$.
	\STATE Update $\psi$ with \textbf{opt}$_A$, $\Delta \psi$ and learning rate $\alpha^A \deltat$.
	\STATE Update $\phi$ with \textbf{opt}$_\pi$, $\Delta \phi$ and learning rate $\alpha^\pi \deltat$.
	\ENDFOR
	\ENDFOR
\end{algorithmic}