### 2019-01-21 paper/icml_drau/policy_improvement.tex

%! TEX root = supplementary.tex
Let $(s_t)_{t\geq 0}$ be a trajectory sampled from $\pi'$ i.e.\ solution of the equation
\begin{equation}
	ds_t / dt = F(s_t, \pi'(s_t))
\end{equation}
with initial condition $s_0 = s$.

Define
\begin{equation}
	B(T) = \int_{t=0}^T \gamma^t r(s_t, \pi'(s_t)) dt + \gamma^T V^\pi(s_T).
\end{equation}
This function if continuously differentiable, and its derivative is
\begin{align}
	\dot{B}(T) &= \gamma^T r(s_T, \pi'(s_T)) + \gamma^T \partial_s V^\pi(s) F(s, \pi'(s)) + \gamma^T \ln(\gamma) V^\pi(s_T)\\
		   &= \gamma^T A^\pi(s_T, \pi'(s))
		   &\geq 0.
\end{align}
Thus $B$ is increasing, and $B(0) = V^\pi(s)$, $\lim\limits_{T\rightarrow \infty} B(t) = V^{\pi'}(s)$.
Consequently, $V^\pi(s) \leq V^{\pi'}(s)$, which ends the proof.