### 2019-01-22 paper/icml_drau/proof_conv_adv.tex

%! TEX root = supplementary.tex
To make things more precise, we denote $\tilde{s}_\deltat^t(s_0)$ the
evaluation at instant $t$ of the solution of $d\tilde{s}_t/dt = F(\tilde{s}_t,
\pi(s_0))$ with starting point $s_0$.

Bellman equation on $Q^\pi_\deltat$ yields
\begin{align}
	Q^\pi_\deltat(s, a) &= r(s, a) \deltat + \gamma^\deltat V^\pi_\deltat(\tilde{s}_\deltat^\deltat(s)).
\end{align}

For all $s$, a first order taylor expansion yields
\begin{equation}
	\tilde{s}_\deltat^\deltat(s) = s + F(s, a)\deltat + \bigO(\deltat^2)
\end{equation}
thus, by uniform boundedness of the hessian of $V^\pi_\deltat$,\footnote{
	Without boundedness of the hessian, we cannot write the second order taylor expansion of
	$V^\pi_\deltat(\tilde{s}_\deltat^\deltat(s))$ in term of $\deltat$.
}
\begin{equation}
	Q^\pi_\deltat(s, a) = r(s, a) \deltat + (1 + \ln(\gamma) \deltat + \bigO(\deltat^2)) (V^\pi_\deltat(s) + \deltat \partial_s V^\pi_\deltat(s) F(s, a) + \bigO(\deltat^2).
\end{equation}
Now, this yields
\begin{equation}
	A^\pi_\deltat(s, a) = r(s, a) + \ln(\gamma) V^\pi_\deltat(s) + \partial_s V^\pi_\deltat(s) F(s, a) + \bigO(\deltat),
\end{equation}
and using the convergence of $V^\pi_\deltat(s)$ to $V^\pi(s)$ (Thm.~\ref{th:conv-value}) and $\partial_s V^\pi_\deltat(s)$ to $\partial_s V^\pi(s)$ (hypothesis) yields the result with 
\begin{equation}
	A^\pi(s, a) = r(s, a) + \ln(\gamma) V^\pi(s) + \partial_s V^\pi(s) F(s, a).
	\label{eq:adv_function}
\end{equation}