### 2019-01-22 paper/icml_drau/experiments.tex

%! TEX root = icml_drau.tex
\section{Experiments}
\label{sec:exp}
\input{lc_fig.tex}
The experiment set provided here is specifically aimed at showing that
the proposed method DAU can operate efficiently on a wide range of time
discretizations, without specific tuning, while standard deep Q-learning
approaches cannot. DAU is compared to DDPG in continuous action settings and to DQN in
discrete action settings. \TODO{(We only test off-policy
algorithms.)}

In all settings, we use the learning algorithm described in
Alg.~\ref{alg:dau} and supplementary, Algo 1. The variants of DDPG and DQN
used are described in the supplementary material, as well as all the hyperparameters
used. To make comparison fair, we tested to different scalings for DDPG and DQN.
One where we use the properly rescaled discount
factor $\gamma_\deltat$ and reward $r_\deltat$ from \eqref{eq:def-gamma},
as well as scaled learning rates. One where we only scaled the discount factor
(to avoid shortsightedness), but left all other hyperaparameters constant
across time discretizations. Experiments on the latter yielded better result
and is presented hereby. Experiments on the first variant are presented in the
supplementary material. \TODO{NDY not
convincing: precisely because DDPG and DQN are not invariant, they might
benefit from a different scaling.}

Let us stress that the quantities plotted are rescaled to make comparison
possible across different timesteps. For example,
return results are given in term of discretized return $R_\deltat$ as defined in \eqref{eq:discretized-return},\footnote{This mostly amounts to scaling rewards
by a factor $\deltat$ when this scaling is not naturally done in the environment. Environment-specific
details are given in the supplementary material.} and, most notably, time elapsed is always given in
normalized epochs (physical time), i.e. the true number of epochs multiplied by the
discretization timestep.

\paragraph{Qualitative experiments: Visualizing policies and values.}
To provide qualitative results, and check robustness to time
discretization both in terms of raw return results and in terms
of convergence of the approximate value function and policies, we first provide results on the simple pendulum environment
from the OpenAI Gym classic control suite.  Its state space is of
dimension $2$. We visualize both the learnt value and policy functions by
plotting, for each point of the phase diagram $(\theta, \dot{\theta})$,
its value and policy. The results are presented in
Fig.~\ref{fig:pend} (policy function) and in Fig.\TODO{ref} in
Supplementary \TODO{ref} (value function).

We plot the learnt policy at several instants in physical time (or
normalized epochs) during training, for various time discretizations
$\deltat$, for both DAU and DDPG. With DAU, the agent's policy quickly
converges for every time discretization without specific hyperparameter
tuning. On the contrary, with DDPG, the agent is only able to learn for
the two largest time discretizations; DDPG learning gets harder as $\deltat$ decreases.

\paragraph{Quantitative experiments.}
We benchmark DAU against DDPG on classic control benchmarks: Pendulum,
Cartpole, BipedalWalker, Ant, and Half-Cheetah environments from OpenAI
Gym. For Pendulum and Ant, DDPG is able to learn for the highest
$\deltat$ but fails for the others, while DAU remains reasonably
invariant to the time discretization. On BipedalWalker, DDPG is never
able to learn while DAU is achieving decent performance for all the
$\deltat$'s. On Cartpole and Half-Cheetah, the results are less clear. On
Cartpole, noise dominates, which makes interpretation difficult. On Half-Cheetah, DAU is not clearly more invariant to time discretization than
DDPG. This could be explained by the multiple suboptimal regimes that
coexist in the Half-Cheetah environment (walking on the head, walking on
the back). \TODO{Removed, was unclear: ``and that DAU may fail to correctly weight the performance of
each regime (as explained in Section \ref{sec:discussions}).'' Anyway
this regimes exist for all algorithms...}

% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\columnwidth]{figs_data/cartpole_lc.png}
%   \caption{Cartpole}
%   \label{fig:cartpole-lc}
% \end{figure}

% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\columnwidth]{figs_data/ant_lc.png}
%   \caption{Ant}
%   \label{fig:ant-lc}
% \end{figure}
% 
% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\columnwidth]{figs_data/pendulum_lc.png}
%   \caption{Pendulum}
%   \label{fig:pendulum-lc}
% \end{figure}

%%% Local Variables:
%%% TeX-master: "icml_drau"
%%% End: