### 2019-01-22 paper/icml_drau/discussions.tex

%! TEX root = icml_drau.tex
\section{Discussion}
\label{sec:discussions}

The method derived in this work seems to yield improved robustness to
time discretization on various environments, e.g. pendulum swing up, or
simple locomotion tasks.  However on certain environments, there is still
room for improvment. In what follows, we identify some of the issues
that could explain this theoretical/practical discrepancy.

\paragraph{Smoothness of the value function.} In all of our proofs, $V^\pi$ is
assumed to be continuously differentiable for all $\pi$.  This %is a strong
hypothesis is not always satisfied in practice. Very simple environments
break this hypothesis, such as the pendulum swing-up environment,
with pendulum height as reward and torque as action.  Indeed, 
% assume that $\pi$
% is the optimal policy for the pendulum swing-up environment, with state
% representation $(\theta, \dot{\theta})$. 
the state space $(\theta, \dot{\theta})$ for this environment
can be partitioned into two subspaces $\mathcal{G}$ and $\mathcal{B}$, with
$\mathcal{G}$ the subspace where the optimal policy 
can swing up the pendulum without crossing the $y$-axis, and $\mathcal{B}$ where
it has to cross the $y$-axis. The difference of $V$-function between a point in
$\mathcal{G}$ and a point in $\mathcal{B}$ is lower bounded by a finite
quantity. 
%Besides, one can find two points $s_1 \in \mathcal{G}$ and $s_2 \in \mathcal{B}$
%such that $s_1$ and $s_2$ are arbitrarily close.
Consequently, the value function is
discontinuous on the frontier between $\mathcal{G}$ and $\mathcal{B}$, which results in non infinitesimal advantages
for actions taken on the frontier. \TODO{Non infinitesimal advantages are found,
more generally, as soon as a given policy has different regimes depending on
where it starts in state space. For such policies the value
function is typically discontinuous at the frontier between two
regimes, resulting in non infinitesimal advantages. The
continuous-time DAU properties would only hold almost-everywhere.}

\paragraph{Memory buffer size scalings}~Thm.~\ref{th:cont-params} is stated for
transitions sampled sequentially from a fixed infinite size sequence. In practice,
transitions are sampled from a memory buffer, to prevent excessive correlations.
The memory buffer used is a fixed size circular buffer, filled
with samples from a single infinite exploratory trajectory. In
our experiments, the same buffer size is used for all time discretization. When
$\deltat$ goes to $0$, this corresponds to learning nearly online on the
infinite sequence, since the buffer will only contain examples that are very
close temporally to the end of the sequence in physical time. \TODO{this
contradicts the use of the buffer against correlations} While the resulting
algorithm converges to a continuous time limit, in the strictest sense it is not
time discretization invariant, as the freshness of samples in the buffer will
vary with the time discretization.
A different choice would be to scale the size of the buffer by a factor of
$\frac{1}{\deltat}$. This would amount to always learning on the same physical time interval of
the trajectory. Our theoretical analysis does not cover this case.
\TODO{a bit too many details here?}

\paragraph{Near continuous time reinforcement learning and RMSProp.} RMSProp~\cite{rmsprop}
estimates a batched version of the second moment of gradients via moving
averages, and divides gradient steps by the square root of this estimate.
This may interact with the learning rate scaling discussed above. In
deterministic environments, gradients typically scale as $\bigO(1)$ in
term of $\deltat$, as seen in \eqref{eq:approx_bellman_A}.  In that case, RMSProp
preconditioning has no effect on the suitable order of magnitude for learning
rates. However, in near continuous \emph{stochastic} environments (Eq.~\ref{eq:sde}), variance of the gradients typically scales as
$\bigO\left(\frac{1}{\deltat}\right)$. With a fixed batch size,
RMSProp will multiply gradients by a factor $\bigO(\sqrt{\deltat})$. In
that case,
learning rates need only be scaled as $\alpha \,\sqrt{\deltat}$ instead of
$\alpha \,\deltat$.

\TODO{+ discuss the use of LSTMs to provide memory/deal with
non-observability; Such LSTMs should take timescales into account, eg
chrono init?}