### 2019-01-23 paper/icml_drau/experiments.tex
.

%! TEX root = icml_drau.tex
\section{Experiments}
\label{sec:exp}
\input{lc_fig.tex}
The experiment set provided here is specifically aimed at showing that
the proposed method DAU can operate efficiently on a wide range of time
discretizations, without specific tuning, while standard deep Q-learning
approaches cannot. DAU is compared to DDPG in continuous action settings and to DQN in
discrete action settings.
We work off-policy; this excludes methods such as \TODO{(A3C, PPO, TRPO, ...)}.

In all settings, we use the learning algorithm described in
Alg.~\ref{alg:dau} and supplementary, Algo 1. The variants of DDPG and DQN
used are described in the supplementary material, as well as all the hyperparameters
used. We tested two different setups for DDPG and DQN.
In one, we scaled the discount factor (to avoid shortsightedness with small $\deltat$), but
left all other hyperparameters constant across time discretizations.
In the other, we used the properly rescaled discount
factor and reward from Eq.~\eqref{eq:def-gamma},
as well as $O(\deltat)$ learning rates for RMSProp.  The former yields slightly
better results, and is presented in the main text. Experiments on the second
variant are presented in the
supplementary material.

Let us stress that the quantities plotted are rescaled to make comparison
possible across different timesteps. For example,
return results are given in term of discretized return $R_\deltat$ as defined in \eqref{eq:discretized-return},\footnote{This mostly amounts to scaling rewards
by a factor $\deltat$ when this scaling is not naturally done in the environment. Environment-specific
details are given in the supplementary material.} and, most notably, time elapsed is always given in
normalized epochs (physical time), i.e. the true number of epochs multiplied by the
discretization timestep.

\paragraph{Qualitative experiments: Visualizing policies and values.}
To provide qualitative results, and check robustness to time
discretization both in terms of raw return results and in terms
of convergence of the approximate value function and policies, we first provide results on the simple pendulum environment
from the OpenAI Gym classic control suite.  Its state space is of
dimension $2$. We visualize both the learnt value and policy functions by
plotting, for each point of the phase diagram $(\theta, \dot{\theta})$,
its value and policy. The results are presented in
Fig.~\ref{fig:pend} (policy function) and in Fig.\TODO{ref} in
Supplementary \TODO{ref} (value function).

We plot the learnt policy at several instants in physical time (or
normalized epochs) during training, for various time discretizations
$\deltat$, for both DAU and DDPG. With DAU, the agent's policy quickly
converges for every time discretization without specific hyperparameter
tuning. On the contrary, with DDPG, the agent is only able to learn for
the two largest time discretizations; DDPG learning gets harder as $\deltat$ decreases.

\paragraph{Quantitative experiments.}
We benchmark DAU against DDPG on classic control benchmarks: Pendulum,
Cartpole, BipedalWalker, Ant, and Half-Cheetah environments from OpenAI
Gym. For Pendulum and Ant, DDPG is able to learn for the highest
$\deltat$ but fails for the others, while DAU remains reasonably
invariant to the time discretization. On BipedalWalker, DDPG is never
able to learn while DAU is achieving decent performance for all the
$\deltat$'s. On Cartpole and Half-Cheetah, the results are less clear. On
Cartpole, noise dominates, which makes interpretation difficult. On Half-Cheetah, DAU is not clearly more invariant to time discretization than
DDPG. This could be explained by the multiple suboptimal regimes that
coexist in the Half-Cheetah environment (walking on the head, walking on
the back).
% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\columnwidth]{figs_data/cartpole_lc.png}
%   \caption{Cartpole}
%   \label{fig:cartpole-lc}
% \end{figure}

% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\columnwidth]{figs_data/ant_lc.png}
%   \caption{Ant}
%   \label{fig:ant-lc}
% \end{figure}
% 
% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\columnwidth]{figs_data/pendulum_lc.png}
%   \caption{Pendulum}
%   \label{fig:pendulum-lc}
% \end{figure}

%%% Local Variables:
%%% TeX-master: "icml_drau"
%%% End: