### 2019-01-23 paper/icml_drau/pendulum_fig.tex

%! TEX root = icml_drau.tex
\newcommand{\iw}{1cm}
\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{figs_data/pendulum_unscaled/pendulum_fig_val_unscaled.pdf}
	\caption{Value functions obtained by DDPG (unscaled version) and
	DAU at different instants in physical time of training on the
	pendulum swing-up environment. Each image represents the learnt
	value function (the $x$-axis is the angle, and the $y$-axis the angular velocity). The lighter the pixel, the higher the value.}
        \label{fig:pend}
\end{figure*}